---
title: "In-class_Ex06"
format: html
editor: visual
---


```{r}
pacman::p_load(sf, sfdep,tmap, plotly, tidyverse)
```
#01 Read in Data
```{r}
hunan <- st_read(dsn = "data/geospatial",
        layer = "Hunan")
```
```{r}
GDPPC <- read_csv("data/aspatial/hunan_GDPPC.csv")
```

#02 Create a Time Series Cube

Space(X,Y) + Time(Z) , only attribute can change (for here is time)

```{r}
GDPPC_st <- spacetime(GDPPC, hunan, #spatial, attribute data
                      .loc_col = "County",   #indicate this is spatial
                      .time_col = "Year")   #indicate this one is time
                                            #cannot use original time cause that's continuous
                                            #convert day to integer or drop time,only keep ymd
```

Check if has been correctly transformed to space time cube
```{r}
is_spacetime_cube(GDPPC_st) 
```

#03 Computing Gi*

##3.1 Deriving the spatial weights
for Gi*, only distance method (inverse or fixed distance(threshold distance)) is applicable. proximity not applicable
```{r}
GDPPC_nb <- GDPPC_st %>%
  activate("geometry") %>%
  mutate(nb = include_self(
          st_contiguity(geometry)),
          wt = st_inverse_distance(nb, 
                                  geometry, 
                                  scale = 1, 
                                  alpha = 1), #distance decay
         .before = 1) %>%   # ?
  set_nbs("nb") %>%     #之后就不能再sort了会乱
  set_wts("wt")
```

```{r}
gi_stars <- GDPPC_nb %>%
  group_by (Year) %>%
  mutate(gi_star= local_gstar_perm(
    GDPPC, nb, wt)) %>%
  tidyr::unnest(gi_star)
```
      

##3.2 Mann-Kendall Test to individual sectors

看单调性monotonic
H0: No monotonic trend
H1: Monotonic trend is present
reject H0 if p value <0.05,
then tau ranges between -1 ~ 1
-1 decreasing series
1 increasing series
```{r}
cbg <- gi_stars %>%
  ungroup() %>%
  filter(County == "Changsha") |>   #new R function ，就是filter， == %>%
  select(County, Year, gi_star)
```

```{r}
#save as a object through ggpolt
# Plot the result using ggplot2
ggplot(data = cbg,
       aes(x = Year,
           y = gi_star)) +
  geom_line() +
  theme_light()
```

##3.3 Interactive Mann-Kendall Plot
make the ggplot interactive
```{r}
p<- ggplot(data = cbg,
           aes(x=Year,
               y= gi_star))+
  geom_line() +
  theme_light()

ggplotly(p)
```

##3.4 Printing MK Test Report
```{r}
library(Kendall)
cbg %>%
  summarise(mk = list(unclass(
    Kendall::MannKendall(gi_star)
  )))%>%
  tidyr::unnest_wider(mk)
```
##3.5 Easier way to replicate 3.1~3.5 MK Test county-wise
tau:看monotonic的-1 单调递减，1单调递增
约接近0correlation约小
```{r}
ehsa <- gi_stars %>%
  group_by(County) %>%
  summarise(mk = list(
    unclass(
      Kendall::MannKendall(gi_star)
    ))) %>%
  tidyr::unnest_wider(mk)

head(ehsa)
```

#04 Performing Emerging Hotspot Analysis
##4.1
```{r}
ehsa <- emerging_hotspot_analysis(
  x = GDPPC_st,
  .var = "GDPPC",
  k = 1,     #现在和前一个时点的比较，类似t和t-1
  nsim = 99  #模拟的次数
)
```

##4.2 Visualising the distribution of EHSA classes
```{r}
ggplot(data = ehsa,
       aes(x = classification))+
  geom_bar()
```

##4.3 Visualising EHSA
```{r}
# Joining hunan and ehsa data
hunan_ehsa <- hunan %>%
  left_join(ehsa, by = join_by(County == location))
```

Oscilating & sporadic 都是有冷热
但Oscilating是有规律和周期的
sporadic 是偶然一下突然冷/热无规律，eg。大部分直接冷，偶发热
```{r}
# Filtering significant results and plotting using tmap
ehsa_sig <- hunan_ehsa %>%
  filter(p_value < 0.05)

tmap_mode("plot")
tm_shape(hunan_ehsa) +
  tm_polygons() +
  tm_borders(alpha = 0.5) +
tm_shape(ehsa_sig) +
  tm_fill("classification") +
  tm_borders(alpha = 0.4)
```

 