---
title: "In-class_Ex05"
format: html
editor: visual
---

```{r}
pacman::p_load(sf, tmap, sfdep, tidyverse)
```

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

```{r}
hunan_GDPPC <- left_join(hunan, hunan2012) %>%
  select(1:4, 7, 15)
```

```{r}
tmap_mode("plot")
tm_shape(hunan_GDPPC) +
  tm_fill("GDPPC", 
          style = "quantile", 
          palette = "Blues",
          title = "GDPPC") +
  tm_layout(main.title = "Distribution of GDP per capita by county, Hunan Province",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2)
```

p8

```{r}
wm_q <- hunan_GDPPC %>%
  mutate(nb = st_contiguity(geometry),
         wt = st_weights(nb,
                         style = "W"),
         .before = 1) 
```

```{r}
wm_q
```

```{r}
moranI <- global_moran(wm_q$GDPPC,
                       wm_q$nb,
                       wm_q$wt)
glimpse(moranI)
```

```{r}
global_moran_test(wm_q$GDPPC,
                       wm_q$nb,
                       wm_q$wt)
```

p12

```{r}
set.seed(1234)
```

```{r}
global_moran_perm(wm_q$GDPPC,
                       wm_q$nb,
                       wm_q$wt,
                  nsim = 99)
```

和handson的区别在于inclass use the new way, hands-on use old way

13 Computing local Moran's l

```{r}
lisa <-wm_q %>%
    mutate(local_moran = local_moran(
      GDPPC, nb, wt, nsim = 99),
        .before = 1) %>%
  unnest(local_moran)
```

```{r}
tmap_mode("plot")
tm_shape(lisa) +
  tm_fill("ii") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(
    main.title = "local Moran's I of GDPPC",
    main.title.size = 2)
```

```{r}
tmap_mode("plot")
tm_shape(lisa) +
  tm_fill("p_ii_sim") + 
  tm_borders(alpha = 0.5) +
   tm_layout(main.title = "p-value of local Moran's I",
            main.title.size = 2)
```

```{r}
tmap_mode("plot")
map1 <- tm_shape(lisa) +
  tm_fill("ii") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(main.title = "local Moran's I of GDPPC",
            main.title.size = 0.8)

map2 <- tm_shape(lisa) +
  tm_fill("p_ii",
          breaks = c(0, 0.001, 0.01, 0.05, 1),
              labels = c("0.001", "0.01", "0.05", "Not sig")) + 
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "p-value of local Moran's I",
            main.title.size = 0.8)

tmap_arrange(map1, map2, ncol = 2)
```

```{r}
lisa_sig <- lisa  %>%
  filter(p_ii_sim < 0.05)
tmap_mode("plot")
tm_shape(lisa) +
  tm_polygons() +
  tm_borders(alpha = 0.5) +
tm_shape(lisa_sig) +
  tm_fill("mean") + 
  tm_borders(alpha = 0.4)
```

```{r}
wm_idw <- hunan_GDPPC %>%
  mutate(nb = include_self(
    st_contiguity(geometry)),
    wts = st_inverse_distance(nb, 
                              geometry, 
                              scale = 1,
                              alpha = 1),
         .before = 1)
```

```{r}
HCSA <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    GDPPC, nb, wts, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)
HCSA
```

```{r}
tmap_mode("plot")
tm_shape(HCSA) +
  tm_fill("gi_star") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8))
```

```{r}
tmap_mode("plot")
tm_shape(HCSA) +
  tm_fill("p_sim") + 
  tm_borders(alpha = 0.5)
```

```{r}
tmap_mode("plot")
map1 <- tm_shape(HCSA) +
  tm_fill("gi_star") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(main.title = "Gi* of GDPPC",
            main.title.size = 0.8)

map2 <- tm_shape(HCSA) +
  tm_fill("p_value",
          breaks = c(0, 0.001, 0.01, 0.05, 1),
              labels = c("0.001", "0.01", "0.05", "Not sig")) + 
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "p-value of Gi*",
            main.title.size = 0.8)

tmap_arrange(map1, map2, ncol = 2)
```

## Global Measures of Spatial Association

### Step 1: Deriving Queen’s contiguity weights: sfdep methods

```{r}
wm_q <- hunan_GDPPC %>%
  mutate(nb = st_contiguity(geometry),
         wt = st_weights(nb,
                         style = "W"),
         .before = 1) 
```

### Computing Global Moran’ I

In the code chunk below, [`global_moran()`](https://sfdep.josiahparry.com/reference/global_moran) function is used to compute the Moran’s I value. Different from spdep package, the output is a tibble data.frame.

```{r}
moranI <- global_moran(wm_q$GDPPC,
                       wm_q$nb,
                       wm_q$wt)
glimpse(moranI)
```

### Performing Global Moran’sI test

In general, Moran’s I test will be performed instead of just computing the Moran’s I statistics. With sfdep package, Moran’s I test can be performed by using [`global_moran_test()`](https://sfdep.josiahparry.com/reference/global_moran_test.html) as shown in the code chunk below.

```{r}
global_moran_test(wm_q$GDPPC,
                       wm_q$nb,
                       wm_q$wt)
```

### Performing Global Moran’I permutation test

In practice, Monte carlo simulation should be used to perform the statistical test. For **sfdep**, it is supported by [`globel_moran_perm()`](https://sfdep.josiahparry.com/reference/global_moran_perm.html)

```{r}
set.seed(1234)
```

Next, `global_moran_perm()` is used to perform Monte Carlo simulation.

```{r}
global_moran_perm(wm_q$GDPPC,
                       wm_q$nb,
                       wm_q$wt,
                  nsim = 99)
```

## LISA map

LISA map is a categorical map showing **outliers** and **clusters**. There are two types of outliers namely: High-Low and Low-High outliers. Likewise, there are two type of clusters namely: High-High and Low-Low cluaters. In fact, LISA map is an interpreted map by combining local Moran’s I of geographical areas and their respective p-values.

## Computing local Moran’s I

In this section, you will learn how to compute Local Moran’s I of GDPPC at county level by using [`local_moran()`](https://sfdep.josiahparry.com/reference/local_moran.html) of sfdep package.

```{r}
lisa <- wm_q %>% 
  mutate(local_moran = local_moran(
    GDPPC, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_moran)
```

### Visualising local Moran’s I

In this code chunk below, tmap functions are used prepare a choropleth map by using value in the *ii* field.

```{r}
tmap_mode("plot")
tm_shape(lisa) +
  tm_fill("ii") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(
    main.title = "local Moran's I of GDPPC",
    main.title.size = 2)
```

### Visualising p-value of local Moran’s I

In the code chunk below, tmap functions are used prepare a choropleth map by using value in the *p_ii_sim* field.

```{r}
tmap_mode("plot")
tm_shape(lisa) +
  tm_fill("p_ii_sim") + 
  tm_borders(alpha = 0.5) +
   tm_layout(main.title = "p-value of local Moran's I",
            main.title.size = 2)
```

### Visualising local Moran’s I and p-value

For effective comparison, it will be better for us to plot both maps next to each other.

```{r}
tmap_mode("plot")
map1 <- tm_shape(lisa) +
  tm_fill("ii") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(main.title = "local Moran's I of GDPPC",
            main.title.size = 0.8)

map2 <- tm_shape(lisa) +
  tm_fill("p_ii",
          breaks = c(0, 0.001, 0.01, 0.05, 1),
              labels = c("0.001", "0.01", "0.05", "Not sig")) + 
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "p-value of local Moran's I",
            main.title.size = 0.8)

tmap_arrange(map1, map2, ncol = 2)
```

### Plotting LISA map

In lisa sf data.frame, we can find three fields contain the LISA categories. They are *mean*, *median* and *pysal*. In general, classification in *mean* will be used as shown in the code chunk below.

```{r}
lisa_sig <- lisa  %>%
  filter(p_ii_sim < 0.05)
tmap_mode("plot")
tm_shape(lisa) +
  tm_polygons() +
  tm_borders(alpha = 0.5) +
tm_shape(lisa_sig) +
  tm_fill("mean") + 
  tm_borders(alpha = 0.4)
```

## Hot Spot and Cold Spot Area Analysis (HCSA)

HCSA uses spatial weights to identify locations of statistically significant hot spots and cold spots in an spatially weighted attribute that are in proximity to one another based on a calculated distance. The analysis groups features when similar high (hot) or low (cold) values are found in a cluster. The polygon features usually represent administration boundaries or a custom grid structure.

### Computing local Gi\* statistics

As usual, we will need to derive a spatial weight matrix before we can compute local Gi\* statistics. Code chunk below will be used to derive a spatial weight matrix by using sfdep functions and tidyverse approach.

```{r}
wm_idw <- hunan_GDPPC %>%
  mutate(nb = include_self(
    st_contiguity(geometry)),
    wts = st_inverse_distance(nb, 
                              geometry, 
                              scale = 1,
                              alpha = 1),
         .before = 1)
```

### Computing local Gi\* statistics

Now, we will compute the local Gi\* by using the code chunk below.

```{r}
HCSA <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    GDPPC, nb, wts, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)
HCSA
```

```{r}
tmap_mode("plot")
tm_shape(HCSA) +
  tm_fill("gi_star") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8))
```

### Visualising p-value of HCSA

In the code chunk below, tmap functions are used to plot the p-values of local Gi\* (i.e. p_sim) at the province level.

```{r}
tmap_mode("plot")
tm_shape(HCSA) +
  tm_fill("p_sim") + 
  tm_borders(alpha = 0.5)
```

### Visuaising local HCSA

For effective comparison, you can plot both maps next to each other as shown below.

```{r}
tmap_mode("plot")
map1 <- tm_shape(HCSA) +
  tm_fill("gi_star") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(main.title = "Gi* of GDPPC",
            main.title.size = 0.8)

map2 <- tm_shape(HCSA) +
  tm_fill("p_value",
          breaks = c(0, 0.001, 0.01, 0.05, 1),
              labels = c("0.001", "0.01", "0.05", "Not sig")) + 
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "p-value of Gi*",
            main.title.size = 0.8)

tmap_arrange(map1, map2, ncol = 2)
```

### Visualising hot spot and cold spot areas

Now, we are ready to plot the significant (i.e. p-values less than 0.05) hot spot and cold spot areas by using appropriate tmap functions as shown below.

```{r}
HCSA_sig <- HCSA  %>%
  filter(p_sim < 0.05)
tmap_mode("plot")
tm_shape(HCSA) +
  tm_polygons() +
  tm_borders(alpha = 0.5) +
tm_shape(HCSA_sig) +
  tm_fill("cluster") + 
  tm_borders(alpha = 0.4)
```

14

```{r}
tmap_mode("plot")
tm_shape(lisa) +
  tm_fill("ii") +
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6, 8)) +
  tm_layout(
    main.title = "Local Moran's I of G",
    main.title.size = 2
  )

```
