---
title: "In-class-Exercise 1"
author: "ZHANG Jinghan"
date: "Aug 20 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  message: false
  freeze: true
---

```         
```

### The code

p3

```{r}
pacman::p_load(sf, tidyverse, tmap, ggstatsplot)
```

p4

```{r}
mpsz14_shp <- st_read(dsn = "data/",
                layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
mpsz14_kml <- st_read("data/MasterPlan2014SubzoneBoundaryWebKML.kml")
```

05

```{r}
st_write(mpsz14_shp, 
         "data/MP14_SUBZONE_WEB_PL.kml",
         delete_dsn = TRUE)
```

06 import Pre-Schools Location in kml geojson save them in sf simple features data frame.

```{r}
preschool_kml <- st_read("data/PreSchoolsLocation.kml")
```

```{r}
preschool_geojson <- st_read("data/PreSchoolsLocation.geojson") 
```

07

```{r}
mpsz19_shp <- st_read(dsn = "data/",
                layer = "MPSZ-2019")
```

```{r}
mpsz19_kml <- st_read("data/MasterPlan2019SubzoneBoundaryNoSeaKML.kml")
```

08 Checking coordinate system of sf project

```{r}
st_crs(mpsz19_shp)
```

09 Transformation Re-write the code chunk to import the Master Plan Sub-zone 2019 and Pre-schools Location with proper transformation

```{r}
mpsz19_shp <- st_read(dsn = "data/",
                layer = "MPSZ-2019") %>%
  st_transform(crs = 3414)
```

```{r}
preschool <- st_read("data/PreSchoolsLocation.kml") %>%
  st_transform(crs = 3414)
```

##Geospatial Data Wragling 10 Point-in-Polygon count count the number of pre-schools in each planning sub-zone.

```{r}
mpsz19_shp <- mpsz19_shp %>%
  mutate(`PreSch Count` = lengths(
    st_intersects(mpsz19_shp, preschool)))
```

11 Compute density

```{r}
mpsz19_shp <- mpsz19_shp %>%
  mutate(Area = units::drop_units(
    st_area(.)),
    `PreSch Density` = `PreSch Count` / Area * 1000000
  )
```

##Statistical Analysis 12 EDA and CDA Confirmatory Data Analysis Tip: Refer to ggscatterstats() of ggstatsplot package. ###\`\``{r} mpsz$`PreSch Density`<- as.numeric(as.character(mpsz19_shp$`PreSch Density`)) mpsz$`PreSch Count`<- as.numeric(as.character(mpsz19_shp$`PreSch Count\`)) mpsz19_shp \<- as.data.frame(mpsz19_shp)

ggscatterstats(data = mpsz19_shp, x = `PreSch Density`, y = `PreSch Count`, type = "parametric") ###\`\`\`

```{r}
popdata <- read_csv("data/respopagesextod2023.csv")
```

#! output: false st_write(mpsz14_shp, "data/MasterPlan2014SubzoneBoundaryWebKML.kml", delete_dsn = TRUE)

```{r}
mpsz19_kml = st_read("C:/ISSS626-AY2024-25Aug/data/geospatial/MasterPlan2019SubzoneBoundaryNoSeaKML.kml")
```

```{r}
mpsz19_shp = st_read("C:/ISSS626-AY2024-25Aug/data/geospatial/MasterPlan2019SubzoneBoundaryNoSeaGEOJSON.geojson")
```

```{r}
mpsz19_shp <- st_read(dsn = "C:/ISSS626-AY2024-25Aug/data/geospatial",
                      layer = "MPSZ-2019")%>%
  st_transform(crs=3414)
```
